A Constraint Solving Approach to Parikh Images of Regular Languages

### 2022-07-08 main.tex
Fix fudge
cleanup

In the derivation tree of \cref{fig:derivation:multi} we start as before, but
with the product version of the $\Image$ predicate. The only possible rule here
is $\ExpandM$, which we use to add the corresponding constraints on each
automata of the product as we would have had in the single-automaton version. We
perform algebraic reasoning on those equations, along with the constraint on
$\MonoidElement$ to determine bounds on transition variables that will enable us
to \Subsume{} and remove one of the $\Connected{}$ predicates. When we have
finished doing so, we use \Materialise{} to compute the product of the now
filtered automata. Further algebraic reasoning allows us to remove the final
instance of $\Connected$. We then perform \Expand{} to get rid of the final
$\Image$ instance, and are immediately able to $\Subsume$ the resulting
$\Connected{}$ predicate. This, again, leaves us with a set of linear
inequalities that we can use to obtain a model of our final $\MonoidElement$
values (now a vector). Performing quantifier elimination on the remaining clauses would also produce the Presburger formula of the Parikh image under the constraint we started with.

@@

  \hypo{\bigstar}
  \infer1[]{

@@

Extending \Calculus{} to support fully symbolic automata is possible within the
framework, depending on your choice of $\Map$. The difficulty consists in
handling the mapping of the homomorphism $\Map$ over symbolic labels, assuming
it maps to finitely many monoid elements. This is not always straightforward, as
seen in the example of \cref{sec:multiple:example}. This complexity is inherent
in computing the full Parikh image, and stems from the fact that we need to
differentiate between the possible interpretations of the $\Sigma$ transitions
without knowing ahead of times which ones will actually be materialied in the
product. If, on the other hand, if our $\Map$ had been length-counting, which
does not differentiate between values, it would have required no adaptation at
all. With a somewhat liberal interpretation of what we are allowed to map to
(e.g. fresh terms), it is possible to use $\Map$ to represent choice operations
like the ones in a range label. In \Catra{}, we frontload this problem by
requiring the user to encode their input as a Parikh automaton. \Fudge{How is
the encoding of Ostrich Plus automata done?}

>>>>

In the derivation tree of \cref{fig:derivation:multi} we start as before, but
with the product version of the $\Image$ predicate. The only possible rule here
is $\ExpandM$, which we use to add the corresponding constraints on each
automata of the product as we would have had in the single-automaton version. We
perform algebraic reasoning on those equations, along with the constraint on
$\MonoidElement$ to determine bounds on transition variables that will enable us
to \Subsume{} and remove one of the $\Connected{}$ predicates. When we have
finished doing so, we use \Materialise{} to compute the product of the now
filtered automata. Further algebraic reasoning allows us to remove the final
instance of $\Connected$. We then perform \Expand{} to get rid of the final
$\Image$ instance, and are immediately able to $\Subsume$ the resulting
$\Connected{}$ predicate. This, again, leaves us with a set of linear
inequalities at the leaf that we can use to obtain a model of our final
$\MonoidElement$ values (now a vector).

@@

  \hypo{

@@

Extending \Calculus{} to support fully symbolic automata is possible within the
framework, depending on your choice of $\Map$. The difficulty consists in
handling the mapping of the homomorphism $\Map$ over symbolic labels, assuming
it maps to finitely many monoid elements. This is not always straightforward, as
seen in the example of \cref{sec:multiple:example}. This complexity is inherent
in computing the full Parikh image, and stems from the fact that we need to
differentiate between the possible interpretations of the $\Sigma$ transitions
without knowing ahead of times which ones will actually be materialied in the
product. If, on the other hand, if our $\Map$ had been length-counting, which
does not differentiate between values, it would have required no adaptation at
all. With a somewhat liberal interpretation of what we are allowed to map to
(e.g. fresh terms), it is possible to use $\Map$ to represent choice operations
like the ones in a range label. In \Catra{}, we frontload this problem by
requiring the user to encode their input as a Parikh automaton. For an example of how automata can be encoded, see \cref{sec:parikh-automata}.
